---
title: "Untitled"
author: "Shelly Trigg"
date: "7/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

load libraries
```{r}
library(data.table)
library(gplots)
library(ggplot2)
library(dplyr)
library(broom)
library(RColorBrewer)
library(egg)
```

# Part 1:  Filter methylated regions for regions that have coverage in 75% of samples per group

read in data
```{r}
DMR <- fread("https://gannet.fish.washington.edu/metacarcinus/Salmo_Calig/analyses/20200702/DMR250bp_MCmax25_cov5x_rms_results_collapsed.tsv")

```



loop through table and keep only lines where up to one sample contains NA for % methylation
```{r}
## For ambient DMR table
df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(DMR))){
  t16_s26 <- DMR[i,7:10] #define columns from 16C_26psu
  t16_s32 <- DMR[i,11:14] #define columns from 16C_32psu
  t8_s26 <- DMR[i,15:18] #define columns from 8C_26psu
  t8_s32 <- DMR[i,19:22] #define columns from 8C_32psu
  if(length(which(is.na(t16_s26))) < 2 & length(which(is.na(t16_s32))) < 2 & length(which(is.na(t8_s26))) < 2 & length(which(is.na(t8_s32))) < 2){
  df <- rbind(df,DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}

```

# Part 2: Run group statistics on regions to find regions that are significantly different among groups

Make a unique ID column 
```{r}
#for all ambient sample comparison
df$ID <- paste(df$chr,":",df$start,"-",df$end, sep = "")
df$ID <- gsub("__.*__.*:",":",df$ID)

```

reformat data for calculating group effect
```{r}
#reformat all to long format
DMR_STACKED <- tidyr::gather(df[,7:ncol(df)], "Sample.ID", "perc.meth",1:20)

#simplify sample ID column
DMR_STACKED$Sample.ID <- gsub("methylation_level_","", DMR_STACKED$Sample.ID)

#Create treatment column
DMR_STACKED$Treatment <- gsub("psu_.*","psu", DMR_STACKED$Sample.ID)

#create a temperature column
DMR_STACKED$Temp <- gsub("C_.*","",DMR_STACKED$Sample.ID)
DMR_STACKED$Temp <- gsub("CTRL_","", DMR_STACKED$Temp)

#Create a salinity column
DMR_STACKED$Salinity <- gsub(".*C_","", DMR_STACKED$Sample.ID)
DMR_STACKED$Salinity <- gsub("psu_.*", "", DMR_STACKED$Salinity)

#Create an infestation status column
DMR_STACKED$Lice_status <- ifelse(grepl("CTRL",DMR_STACKED$Sample.ID),"no","yes")
```

plot % meth dist.
```{r}
a <- ggplot(DMR_STACKED) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw() + xlab("fraction methylated CpGs") + theme(text = element_text(size=8),plot.title = element_text(size = 8,face = "bold"),legend.title = element_text(size = 4),legend.text = element_text(size = 4))

```

arc sin sqrt transform the data
```{r}
#arcsin sqrt transformation function
asinTransform <- function(p) { asin(sqrt(p))}

#arcsin transform data 

DMR_STACKED_asin <- DMR_STACKED
DMR_STACKED_asin$perc.meth <- asinTransform(DMR_STACKED_asin$perc.meth)

```

plot distribution of TRANSFORMED % methylation in all DMRs in all samples
```{r}
b <- ggplot(DMR_STACKED_asin) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw() + xlab("transformed fraction methylated CpGs") + theme(text = element_text(size=8),plot.title = element_text(size = 8,face = "bold"),legend.title = element_text(size = 4),legend.text = element_text(size = 4))

jpeg("DMR_mCpG_histograms.jpg", width = 6, height = 6, units = "in", res = 300)
ggarrange(a,b, nrow = 2)
dev.off()
```

## Run anova on TRANSFORMED data to assess group differences for each DMR

**Hypotehsis: there is no effect from treatment**
```{r}
DMR_1way_aov <- DMR_STACKED_asin[which(DMR_STACKED_asin$Lice_status == "yes"),] %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Treatment, data =  . ))
#summarize ANOVA data
DMR_1way_aov_modelsumm <- glance(DMR_1way_aov, meth_aov_models)
```

**Hypotehsis: there is no effect from temp, sal., or their interaction**
```{r}
#
DMR_2way_aov <- DMR_STACKED_asin[which(DMR_STACKED_asin$Lice_status == "yes"),] %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Temp*Salinity, data =  . ))
#summarize ANOVA data
DMR_2way_aov_modelsumm <- glance(DMR_2way_aov, meth_aov_models)
DMR_2way_aov_modelsumm_0.05 <- DMR_2way_aov_modelsumm[which(DMR_2way_aov_modelsumm$p.value <= 0.05),]


#run tukey test 
DMR_2way_aov_tuk <- DMR_STACKED_asin[which(DMR_STACKED_asin$Lice_status == "yes"),] %>% group_by(ID) %>%
do(meth_tuk_models = TukeyHSD(aov(perc.meth~Temp*Salinity, data =  . )))

DMR_2way_aov_tuk_modelsumm <- tidy(DMR_2way_aov_tuk,meth_tuk_models)

#filter tuk results for DMRs with overall model pvalue < = 0.05
DMR_2way_aov0.05_tuk_modelsumm <- DMR_2way_aov_tuk_modelsumm[which(DMR_2way_aov_tuk_modelsumm$ID %in% DMR_2way_aov_modelsumm_0.05$ID),]
#there are 83 DMRs

#filter for DMRs with Tuk p <= 0.05
DMR_2way_aov0.05_tuk_modelsumm_0.05 <- DMR_2way_aov0.05_tuk_modelsumm[which(DMR_2way_aov0.05_tuk_modelsumm$adj.p.value<= 0.05),]
#there are 81 DMRs; 2 DMRs did not have adj.p.values <= 0.05

DMR_2way_aov0.05_tuk_modelsumm_0.05$term_comp <- paste(DMR_2way_aov0.05_tuk_modelsumm_0.05$term, DMR_2way_aov0.05_tuk_modelsumm_0.05$comparison, sep = "_")


DMR_2way_aov_tuk_modelsumm_0.05_spread <- data.frame(tidyr::spread(DMR_2way_aov0.05_tuk_modelsumm_0.05[,c(1,7,8)], "term_comp", "adj.p.value"))

#add column for chr, start and end

DMR_2way_aov_tuk_modelsumm_0.05_spread$chr <- gsub(":.*","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)

DMR_2way_aov_tuk_modelsumm_0.05_spread$start <- gsub(".*:","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)

DMR_2way_aov_tuk_modelsumm_0.05_spread$start <- gsub("-.*","",DMR_2way_aov_tuk_modelsumm_0.05_spread$start)

DMR_2way_aov_tuk_modelsumm_0.05_spread$end <- gsub(".*-","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)

DMR_2way_aov_tuk_modelsumm_0.05_spread_bed <- DMR_2way_aov_tuk_modelsumm_0.05_spread[,c(10:12, 2:9)]

write.table(DMR_2way_aov_tuk_modelsumm_0.05_spread_bed,"DMR_2way_aov_tuk_modelsumm_0.05.bed", col.names = F, row.names = F, quote = F)
```

**Hypotehsis: there is no effect from temp, sal.,lice, or their interactions**
This does not work well because there is not enough data for the CTRLs 
So there is no output for Salinity x Lice_status or the three way interaction


```{r}
#
DMR_3way_aov <- DMR_STACKED_asin %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Temp*Salinity*Lice_status, data =  . ))
#summarize ANOVA data
DMR_3way_aov_modelsumm <- tidy(DMR_3way_aov, meth_aov_models)


#I also tried 
# library(rstatix)
# anova_test(perc.meth~Temp*Salinity*Lice_status, data =  DMR_STACKED_asin )
#Coefficient covariances computed by hccm()
#Note: model has aliased coefficients
#      sums of squares computed by model comparison
#ANOVA Table (type II tests)

#                     Effect DFn   DFd     F     p p<.05      ges
#1                      Temp   1 14467 0.273 0.601       1.89e-05
#2                  Salinity   1 14467 2.033 0.154       1.41e-04
#3               Lice_status   1 14467 3.516 0.061       2.43e-04
#4             Temp:Salinity   1 14467 2.790 0.095       1.93e-04
#5          Temp:Lice_status   1 14467 0.025 0.875       1.72e-06
#6      Salinity:Lice_status   0 14467    NA    NA  <NA>       NA
#7 Temp:Salinity:Lice_status   0 14467    NA    NA  <NA>       NA
```

write out ANOVA data to file
```{r}
#write ANOVA output to file
write.csv(rbind(d10DMR_1way_aov_pH_modelsumm,d135DMR_1way_aov_pH_modelsumm,d145DMR_1way_aov_pH_modelsumm,ambDMR_1way_aov_pH_modelsumm), "RAnalysis/Output/DMR_1way_aov_modelsumm.csv", row.names = FALSE, quote = FALSE)
```

plot heatmap of 1way aov for pH p.val 0.05 sig DMRs
```{r}
#create matrix for samples
aov1way_0.05_DMR_m <- as.matrix(df[,7:22])
rownames(aov1way_0.05_DMR_m) <- df$ID

aov1way_0.05_DMR_m <- aov1way_0.05_DMR_m[which(rownames(aov1way_0.05_DMR_m) %in% pull(DMR_1way_aov_modelsumm[which(DMR_1way_aov_modelsumm$p.value < 0.05),],ID)),]
#remove extra text from column names
colnames(aov1way_0.05_DMR_m) <- gsub("methylation_level_","",colnames(aov1way_0.05_DMR_m))

heatmap.2(aov1way_0.05_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov_0.05_DMR_m),rowsep=1:nrow(aov_0.05_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))
```

plot heatmap of 2way aov for pH p.val 0.05 sig DMRs
```{r}
#create matrix for looking at salinity effect
aov2way_0.05_Sal_DMR_m <- as.matrix(df[,7:26])
rownames(aov2way_0.05_Sal_DMR_m) <- df$ID

aov2way_0.05_Sal_DMR_m <- aov2way_0.05_Sal_DMR_m[which(rownames(aov2way_0.05_Sal_DMR_m) %in% pull(DMR_2way_aov_tuk_modelsumm_0.05_spread[which(DMR_2way_aov_tuk_modelsumm_0.05_spread$Salinity_32.26 < 0.05),],ID)),]
#remove extra text from column names
colnames(aov2way_0.05_Sal_DMR_m) <- gsub("methylation_level_","",colnames(aov2way_0.05_Sal_DMR_m))

heatmap.2(aov2way_0.05_Sal_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_Sal_DMR_m),rowsep=1:nrow(aov2way_0.05_Sal_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))


#create matrix for looking at temp effect
aov2way_0.05_temp_DMR_m <- as.matrix(df[,7:22])
rownames(aov2way_0.05_temp_DMR_m) <- df$ID

aov2way_0.05_temp_DMR_m <- aov2way_0.05_temp_DMR_m[which(rownames(aov2way_0.05_temp_DMR_m) %in% pull(DMR_2way_aov_tuk_modelsumm_0.05_spread[which(DMR_2way_aov_tuk_modelsumm_0.05_spread$`Temp_8-16` < 0.05),],ID)),]
#remove extra text from column names
colnames(aov2way_0.05_temp_DMR_m) <- gsub("methylation_level_","",colnames(aov2way_0.05_temp_DMR_m))

heatmap.2(aov2way_0.05_temp_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_temp_DMR_m),rowsep=1:nrow(aov2way_0.05_temp_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))
```

#Create dfs for exclusively salinity effect, temp effect, and interaction effect
```{r}
DMR_2way_aov_0.05_temp <- data.frame()

for(i in 1:nrow(DMR_2way_aov_tuk_modelsumm_0.05_spread)){
  rsum <- rowSums(DMR_2way_aov_tuk_modelsumm_0.05_spread[i,2:9], na.rm = T)
  val <- DMR_2way_aov_tuk_modelsumm_0.05_spread[i,3]
  if(!(is.na(val)) & val == rsum){
   df_r <- DMR_2way_aov_tuk_modelsumm_0.05_spread[i,] 
   DMR_2way_aov_0.05_temp <- rbind(DMR_2way_aov_0.05_temp, df_r)
  }
}


DMR_2way_aov_0.05_sal <- data.frame()

for(i in 1:nrow(DMR_2way_aov_tuk_modelsumm_0.05_spread)){
  rsum <- rowSums(DMR_2way_aov_tuk_modelsumm_0.05_spread[i,2:9], na.rm = T)
  val <- DMR_2way_aov_tuk_modelsumm_0.05_spread[i,2]
  if(!(is.na(val)) & val == rsum){
   df_r <- DMR_2way_aov_tuk_modelsumm_0.05_spread[i,] 
   DMR_2way_aov_0.05_sal <- rbind(DMR_2way_aov_0.05_sal, df_r)
  }
}

DMR_2way_aov_0.05_salxtemp <- DMR_2way_aov_tuk_modelsumm_0.05_spread[which(!(DMR_2way_aov_tuk_modelsumm_0.05_spread$ID %in% DMR_2way_aov_0.05_sal$ID) & !(DMR_2way_aov_tuk_modelsumm_0.05_spread$ID %in% DMR_2way_aov_0.05_temp$ID)),]
  
  

#create matrix for heatmap plotting
aov2way_0.05_DMR_m <- as.matrix(df[,7:22])
rownames(aov2way_0.05_DMR_m) <- df$ID

#remove extra text from column names
colnames(aov2way_0.05_DMR_m) <- gsub("methylation_level_","",colnames(aov2way_0.05_DMR_m))


#plot DMRs that show temerature effect @ p < 0.05
aov2way_0.05_temp_DMR_m <- aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% DMR_2way_aov_0.05_temp$ID),]

heatmap.2(aov2way_0.05_temp_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_temp_DMR_m),rowsep=1:nrow(aov2way_0.05_temp_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))

#plot DMRs that show salinity effect @ p < 0.05
aov2way_0.05_sal_DMR_m <- aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% DMR_2way_aov_0.05_sal$ID),]

heatmap.2(aov2way_0.05_sal_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_sal_DMR_m),rowsep=1:nrow(aov2way_0.05_sal_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))

#plot DMRs that show interaction effect @ p < 0.05
aov2way_0.05_salxtemp_DMR_m <- aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% DMR_2way_aov_0.05_salxtemp$ID),]

heatmap.2(aov2way_0.05_salxtemp_DMR_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_salxtemp_DMR_m),rowsep=1:nrow(aov2way_0.05_salxtemp_DMR_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))

```



###Visualize group means
```{r}
#calculate group means
MeanT16S26 <- rowMeans(aov2way_0.05_DMR_m[,1:4], na.rm = TRUE)
MeanT16S32 <- rowMeans(aov2way_0.05_DMR_m[,5:8], na.rm = TRUE)
MeanT8S26 <- rowMeans(aov2way_0.05_DMR_m[,9:12], na.rm = TRUE)
MeanT8S32 <- rowMeans(aov2way_0.05_DMR_m[,13:16], na.rm = TRUE)


#bind all group means together
aov2way_0.05_DMR_mean_m <- as.matrix(data.frame(cbind(MeanT16S26, MeanT16S32, MeanT8S26, MeanT8S32)))

#plot for day10 group mean comparison
colnames(aov2way_0.05_DMR_mean_m) <- c("16C_26psu", "16C_32psu", "8C_26psu", "8C_32psu")


#plot DMRs that show temerature effect @ p < 0.05
aov2way_0.05_temp_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% DMR_2way_aov_0.05_temp$ID),]

jpeg("DMR_T_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov2way_0.05_temp_DMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_temp_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_temp_DMR_mean_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))
dev.off()

#plot DMRs that show salinity effect @ p < 0.05
aov2way_0.05_sal_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% DMR_2way_aov_0.05_sal$ID),]

jpeg("DMR_S_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov2way_0.05_sal_DMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_sal_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_sal_DMR_mean_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))
dev.off()

#plot DMRs that show interaction effect @ p < 0.05
aov2way_0.05_salxtemp_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% DMR_2way_aov_0.05_salxtemp$ID),]

jpeg("DMR_TxS_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov2way_0.05_salxtemp_DMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_salxtemp_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_salxtemp_DMR_mean_m),keysize=0.5, key.par = list(cex=1),lhei=c(1.5,4,1), lwid = c(1.5,4),lmat = rbind(c(0,3),c(2,1),c(4,0)))
dev.off()
```

```

Write out bedfiles
```{r}
#write out d10 DMR bed file
aov_0.05pH_d10DMR_bed <- data.frame(aov_0.05pH_d10DMR_m)
aov_0.05pH_d10DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$start <- gsub("-.*","", aov_0.05pH_d10DMR_bed$start)
aov_0.05pH_d10DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$comparison <- "d10"
aov_0.05pH_d10DMR_bed <- aov_0.05pH_d10DMR_bed[,c((ncol(aov_0.05pH_d10DMR_bed)-3):ncol(aov_0.05pH_d10DMR_bed)) ]
write.table(aov_0.05pH_d10DMR_bed,"RAnalysis/Output/aov_0.05pH_d10DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")

#write out d135 DMR bed file
aov_0.05pH_d135DMR_bed <- data.frame(aov_0.05pH_d135DMR_m)
aov_0.05pH_d135DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$start <- gsub("-.*","", aov_0.05pH_d135DMR_bed$start)
aov_0.05pH_d135DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$comparison <- "d135"
aov_0.05pH_d135DMR_bed <- aov_0.05pH_d135DMR_bed[,c((ncol(aov_0.05pH_d135DMR_bed)-3):ncol(aov_0.05pH_d135DMR_bed)) ]
write.table(aov_0.05pH_d135DMR_bed,"RAnalysis/Output/aov_0.05pH_d135DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")

#write out d145 DMR bed file
aov_0.05pH_d145DMR_bed <- data.frame(aov_0.05pH_d145DMR_m)
aov_0.05pH_d145DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$start <- gsub("-.*","", aov_0.05pH_d145DMR_bed$start)
aov_0.05pH_d145DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$comparison <- "d145"
aov_0.05pH_d145DMR_bed <- aov_0.05pH_d145DMR_bed[,c((ncol(aov_0.05pH_d145DMR_bed)-3):ncol(aov_0.05pH_d145DMR_bed)) ]
write.table(aov_0.05pH_d145DMR_bed,"RAnalysis/Output/aov_0.05pH_d145DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")


#write out amb DMR bed file
aov_0.05pH_ambDMR_bed <- data.frame(aov_0.05pH_ambDMR_m)
aov_0.05pH_ambDMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$start <- gsub("-.*","", aov_0.05pH_ambDMR_bed$start)
aov_0.05pH_ambDMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$comparison <- "time"
aov_0.05pH_ambDMR_bed <- aov_0.05pH_ambDMR_bed[,c((ncol(aov_0.05pH_ambDMR_bed)-3):ncol(aov_0.05pH_ambDMR_bed)) ]
write.table(aov_0.05pH_ambDMR_bed,"RAnalysis/Output/aov_0.05pH_ambDMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")
```



Create boxplots of significant DMR sizes (not very interesting)
```{r}
#bind together all DMR bed files
aov_0.05_DMR_bed <- rbind(aov_0.05pH_ambDMR_bed, aov_0.05pH_d10DMR_bed,aov_0.05pH_d135DMR_bed,aov_0.05pH_d145DMR_bed)

#replace "d" with "day" in the comparison column
aov_0.05_DMR_bed$comparison <- gsub("d","day ",aov_0.05_DMR_bed$comparison)

#create a column for DMR length
aov_0.05_DMR_bed$length <- (as.numeric(aov_0.05_DMR_bed$end) - as.numeric(aov_0.05_DMR_bed$start))

#make violin plots
jpeg("RAnalysis/Output/DMR_size_distributions.jpg", width = 6, height = 4.5, units = "in", res = 300)
ggplot(aov_0.05_DMR_bed,aes(comparison,length, fill = comparison)) + geom_violin(trim = FALSE) + geom_boxplot(width=0.1) + theme_bw() + theme(axis.text = element_text(size = 10))
dev.off()

#transform to normalize
ggplot(aov_0.05_DMR_bed,aes(log(length,2), fill = comparison, color = comparison)) + geom_density(alpha = 0.25)

#test if a pH or time ever has an affect on DMR length
summary(aov(log(length,2)~comparison, data =  aov_0.05_DMR_bed ))


#plot distribution of DMRs across scaffolds
jpeg("RAnalysis/Output/DMR_distribution_across_scaffolds.jpg", width = 10, height = 5, units = "in", res = 300)
ggplot(aov_0.05_DMR_bed, aes(substr(Scaffold,10,11), group = comparison, color = comparison, fill = comparison)) + geom_bar(aes(y = (..prop..)*100), stat="count", position = "dodge") + xlab("Scaffold") + ylab("% DMRs") + theme_bw()
dev.off()

#read in file with scaffold lengths
pgen_scaffolds <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/GENOMES/Panopea-generosa-v1.0.fa.fai", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

#remove extraneous info
pgen_scaffolds <- pgen_scaffolds[,c(1,2)]
#rename columns
colnames(pgen_scaffolds) <- c("Scaffold", "total_scaff_length")
#add scaffold lengths to DMR table
aov_0.05_DMR_bed <- merge(aov_0.05_DMR_bed,pgen_scaffolds, by = "Scaffold")
#create a scaffold position column
aov_0.05_DMR_bed$Scaffold_position <- (as.numeric(aov_0.05_DMR_bed$start)/as.numeric(aov_0.05_DMR_bed$total_scaff_length))

#plot histograms showing the number of DMRs x scaffold position for each comparison facetted by scaffold
jpeg("RAnalysis/Output/DMR_distribution_by_scaffold_position.jpg", width = 6, height = 7, units = "in", res = 300)
ggplot(aov_0.05_DMR_bed) + geom_histogram(aes(Scaffold_position, fill = comparison), position = "dodge", bins = 10) + facet_wrap(~Scaffold, nrow = 5) + xlab("relative scaffold position") + ylab("number of DMRs") + theme_bw() + theme(axis.text.x = element_text(size = 8, angle = 60, hjust= 1),axis.text.y = element_text(size = 10))
dev.off()

```



# Genomic Feature Analysis




read in data
```{r}
# read in features that DMRs overlap with
amb_DMR_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/aov_0.05pH_amb_0408.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

d10_DMR_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/aov_0.05pH_d10_0408.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

d135_DMR_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/aov_0.05pH_d135_0408.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

d145_DMR_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/aov_0.05pH_d145_0408.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

# read in features that covered regions (background) overlap with
amb_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/amb_features.3CpG.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

d10_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/day10_features.3CpG.txt", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

d135_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/day135_features.3CpG.txt", header = FALSE,sep = "\t", stringsAsFactors = FALSE)

d145_feat <- fread("https://gannet.fish.washington.edu/metacarcinus/Pgenerosa/analyses/20200408/day145_features.3CpG.txt", header = FALSE,sep = "\t", stringsAsFactors = FALSE)
```


format data for plotting
```{r}
# ambient data 

amb_DMR_feat <- amb_DMR_feat[,-c(1:3)]
d10_DMR_feat <- d10_DMR_feat[,-c(1:3)]
d135_DMR_feat <- d135_DMR_feat[,-c(1:3)]
d145_DMR_feat <- d145_DMR_feat[,-c(1:3)]

DMR_feat <- rbind(amb_DMR_feat,d10_DMR_feat,d135_DMR_feat,d145_DMR_feat)

colnames(DMR_feat) <- c("comparison", "scaffold", "start", "end", "feature")

DMR_feat$comparison <- gsub("amb","time", DMR_feat$comparison)
DMR_feat$comparison <- gsub("d", "day ", DMR_feat$comparison)

amb_feat$comparison <- "time"
d10_feat$comparison <- "day 10"
d135_feat$comparison <- "day 135"
d145_feat$comparison <- "day 145"

feat <- rbind(amb_feat,d10_feat, d135_feat, d145_feat)

colnames(feat) <- c("scaffold", "start", "end", "feature","comparison")

feat <- feat[,c(5,1:4)]
  
DMR_feat$region <- "DMRs"
feat$region <- "all.regions"

#combine all features and DMRs
feat_combined <- rbind(DMR_feat, feat)
feat_combined$feature <- gsub("3prime_UTR", "put.3primeUTR",feat_combined$feature)
feat_combined$feature <- gsub("put_promoter", "put.promoter",feat_combined$feature)
feat_combined$feature <- gsub("repeat_region", "repeat.reagion",feat_combined$feature)


feat_combined_noTime <- feat_combined[which(feat_combined$comparison!="time"),]

feat_combined_TIME <- feat_combined[which(feat_combined$comparison=="time"),]

#plot all pH comparions
jpeg("RAnalysis/Output/pH.DMR_genom_feat_stacked_barplot.jpg", width = 6, height = 5, units = "in", res =300)
ggplot(feat_combined_noTime,aes(x = region, y = ..count.., fill = factor(feature), group = feature)) + geom_bar(position = "fill", color = "black") + scale_y_continuous(labels=scales::percent) + ylab("% of total regions") + facet_wrap(~comparison,ncol = 4) + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + scale_fill_manual("Feature",values = RColorBrewer::brewer.pal(9,"BuGn"))
dev.off()

#plot only time comparison
jpeg("RAnalysis/Output/Time.DMR_genom_feat_stacked_barplot.jpg", width = 4, height = 7, units = "in", res =300)
ggplot(feat_combined_TIME,aes(x = region, y = ..count.., fill = factor(feature), group = feature)) + geom_bar(position = "fill", color = "black") + scale_y_continuous(labels=scales::percent) + ylab("% of total regions") + facet_wrap(~comparison,ncol = 4) + theme_bw() + theme(axis.text.x = element_text(size = 7, angle = 60, hjust = 1), axis.title = element_text(size = 12, face = "bold")) + scale_fill_manual("Feature",values = RColorBrewer::brewer.pal(9,"BuGn"))
dev.off()


```


# Run Chi square test on feature proportions


```{r}
# create table with feature totals for DMRs
amb_DMR_feat_summary <- data.frame(table(amb_DMR_feat$V8))
#rename columns
colnames(amb_DMR_feat_summary) <- c("feature", "numDMR")
 
# create table with feature totals for background regions
amb_feat_summary <- data.frame(table(amb_feat$V4))
#rename columns
colnames(amb_feat_summary) <- c("feature", "numRegion")

#merge background and DMR feature totals 
amb_feat_summ <- merge(amb_DMR_feat_summary, amb_feat_summary, by = "feature", all = TRUE)

#replace NAs with 0
amb_feat_summ[is.na(amb_feat_summ)] <- 0

#create variables for the sum of all DMR and background features
amb_feat_DMR_total <- sum(amb_feat_summ$numDMR)
amb_feat_region_total <- sum(amb_feat_summ$numRegion)

#create a table with Chi stats for each feature
chi_table <- data.frame() #create empty df
for(i in 1:nrow(amb_feat_summ)){ #loop through each feature
  numBKGDReg <-amb_feat_summ$numRegion[i] # variable for number of background regions for specific feature
  numDMRreg <- amb_feat_summ$numDMR[i] # variable for number of DMRs for specific feature
  totDMR <- (amb_feat_DMR_total - numDMRreg) # variable for total number background regions in all features minus the number of background regions for specific feature
  totBKGD <- (amb_feat_region_total - numBKGDReg)  # variable for total number DMRs in all features minus the number of DMRs for specific feature
  ct <- matrix(c(numBKGDReg,numDMRreg,totBKGD,totDMR), ncol = 2) # create contingency table
  colnames(ct) <- c(as.character(amb_feat_summ$feature[i]),paste0("Not",amb_feat_summ$feature[i])) #label columns of contingency table
  rownames(ct) <- c("Region", "DMR") #lab rows of contingency table
  print(ct)
  x <-  data.frame(broom::tidy(prop.test(ct, correct = FALSE))) # create data frame storing the Chi sq stats results
  x$feature <- as.character(amb_feat_summ$feature[i]) # add feature name to chi sq df
  chi_table <- rbind(chi_table,x) #add chi sq stats to master table
}

#adjust chi pvalues
chi_table$p.adj <- p.adjust(chi_table$p.value, method = "fdr")

#merge chi table with region counts
chi_table <- merge(amb_feat_summ,chi_table, by = "feature")

amb_chi_table <- chi_table
amb_chi_table$comparison <- "amb"
```

day 10
```{r}
# create table with feature totals for DMRs
d10_DMR_feat_summary <- data.frame(table(d10_DMR_feat$V8))
#rename columns
colnames(d10_DMR_feat_summary) <- c("feature", "numDMR")
 
# create table with feature totals for background regions
d10_feat_summary <- data.frame(table(d10_feat$V4))
#rename columns
colnames(d10_feat_summary) <- c("feature", "numRegion")

#merge background and DMR feature totals 
d10_feat_summ <- merge(d10_DMR_feat_summary, d10_feat_summary, by = "feature", all = TRUE)

#replace NAs with 0
d10_feat_summ[is.na(d10_feat_summ)] <- 0

#create variables for the sum of all DMR and background features
d10_feat_DMR_total <- sum(d10_feat_summ$numDMR)
d10_feat_region_total <- sum(d10_feat_summ$numRegion)

#create a table with Chi stats for each feature
chi_table <- data.frame() #create empty df
for(i in 1:nrow(d10_feat_summ)){ #loop through each feature
  numBKGDReg <-d10_feat_summ$numRegion[i] # variable for number of background regions for specific feature
  numDMRreg <- d10_feat_summ$numDMR[i] # variable for number of DMRs for specific feature
  totDMR <- (d10_feat_DMR_total - numDMRreg) # variable for total number background regions in all features minus the number of background regions for specific feature
  totBKGD <- (d10_feat_region_total - numBKGDReg)  # variable for total number DMRs in all features minus the number of DMRs for specific feature
  ct <- matrix(c(numBKGDReg,numDMRreg,totBKGD,totDMR), ncol = 2) # create contingency table
  colnames(ct) <- c(as.character(d10_feat_summ$feature[i]),paste0("Not",d10_feat_summ$feature[i])) #label columns of contingency table
  rownames(ct) <- c("Region", "DMR") #lab rows of contingency table
  print(ct)
  x <-  data.frame(broom::tidy(prop.test(ct, correct = FALSE))) # create data frame storing the Chi sq stats results
  x$feature <- as.character(d10_feat_summ$feature[i]) # add feature name to chi sq df
  chi_table <- rbind(chi_table,x) #add chi sq stats to master table
}

#adjust chi pvalues
chi_table$p.adj <- p.adjust(chi_table$p.value, method = "fdr")

#merge chi table with region counts
chi_table <- merge(d10_feat_summ,chi_table, by = "feature")

d10_chi_table <- chi_table
d10_chi_table$comparison <- "d10"
```

day 135
```{r}
# create table with feature totals for DMRs
d135_DMR_feat_summary <- data.frame(table(d135_DMR_feat$V8))
#rename columns
colnames(d135_DMR_feat_summary) <- c("feature", "numDMR")
 
# create table with feature totals for background regions
d135_feat_summary <- data.frame(table(d135_feat$V4))
#rename columns
colnames(d135_feat_summary) <- c("feature", "numRegion")

#merge background and DMR feature totals 
d135_feat_summ <- merge(d135_DMR_feat_summary, d135_feat_summary, by = "feature", all = TRUE)

#replace NAs with 0
d135_feat_summ[is.na(d135_feat_summ)] <- 0

#create variables for the sum of all DMR and background features
d135_feat_DMR_total <- sum(d135_feat_summ$numDMR)
d135_feat_region_total <- sum(d135_feat_summ$numRegion)

#create a table with Chi stats for each feature
chi_table <- data.frame() #create empty df
for(i in 1:nrow(d135_feat_summ)){ #loop through each feature
  numBKGDReg <-d135_feat_summ$numRegion[i] # variable for number of background regions for specific feature
  numDMRreg <- d135_feat_summ$numDMR[i] # variable for number of DMRs for specific feature
  totDMR <- (d135_feat_DMR_total - numDMRreg) # variable for total number background regions in all features minus the number of background regions for specific feature
  totBKGD <- (d135_feat_region_total - numBKGDReg)  # variable for total number DMRs in all features minus the number of DMRs for specific feature
  ct <- matrix(c(numBKGDReg,numDMRreg,totBKGD,totDMR), ncol = 2) # create contingency table
  colnames(ct) <- c(as.character(d135_feat_summ$feature[i]),paste0("Not",d135_feat_summ$feature[i])) #label columns of contingency table
  rownames(ct) <- c("Region", "DMR") #lab rows of contingency table
  print(ct)
  x <-  data.frame(broom::tidy(prop.test(ct, correct = FALSE))) # create data frame storing the Chi sq stats results
  x$feature <- as.character(d135_feat_summ$feature[i]) # add feature name to chi sq df
  chi_table <- rbind(chi_table,x) #add chi sq stats to master table
}

#adjust chi pvalues
chi_table$p.adj <- p.adjust(chi_table$p.value, method = "fdr")

#merge chi table with region counts
chi_table <- merge(d135_feat_summ,chi_table, by = "feature")

d135_chi_table <- chi_table
d135_chi_table$comparison <- "d135"
```

day 145
```{r}
# create table with feature totals for DMRs
d145_DMR_feat_summary <- data.frame(table(d145_DMR_feat$V8))
#rename columns
colnames(d145_DMR_feat_summary) <- c("feature", "numDMR")
 
# create table with feature totals for background regions
d145_feat_summary <- data.frame(table(d145_feat$V4))
#rename columns
colnames(d145_feat_summary) <- c("feature", "numRegion")

#merge background and DMR feature totals 
d145_feat_summ <- merge(d145_DMR_feat_summary, d145_feat_summary, by = "feature", all = TRUE)

#replace NAs with 0
d145_feat_summ[is.na(d145_feat_summ)] <- 0

#create variables for the sum of all DMR and background features
d145_feat_DMR_total <- sum(d145_feat_summ$numDMR)
d145_feat_region_total <- sum(d145_feat_summ$numRegion)

#create a table with Chi stats for each feature
chi_table <- data.frame() #create empty df
for(i in 1:nrow(d145_feat_summ)){ #loop through each feature
  numBKGDReg <-d145_feat_summ$numRegion[i] # variable for number of background regions for specific feature
  numDMRreg <- d145_feat_summ$numDMR[i] # variable for number of DMRs for specific feature
  totDMR <- (d145_feat_DMR_total - numDMRreg) # variable for total number background regions in all features minus the number of background regions for specific feature
  totBKGD <- (d145_feat_region_total - numBKGDReg)  # variable for total number DMRs in all features minus the number of DMRs for specific feature
  ct <- matrix(c(numBKGDReg,numDMRreg,totBKGD,totDMR), ncol = 2) # create contingency table
  colnames(ct) <- c(as.character(d145_feat_summ$feature[i]),paste0("Not",d145_feat_summ$feature[i])) #label columns of contingency table
  rownames(ct) <- c("Region", "DMR") #lab rows of contingency table
  print(ct)
  x <-  data.frame(broom::tidy(prop.test(ct, correct = FALSE))) # create data frame storing the Chi sq stats results
  x$feature <- as.character(d145_feat_summ$feature[i]) # add feature name to chi sq df
  chi_table <- rbind(chi_table,x) #add chi sq stats to master table
}

#adjust chi pvalues
chi_table$p.adj <- p.adjust(chi_table$p.value, method = "fdr")

#merge chi table with region counts
chi_table <- merge(d145_feat_summ,chi_table, by = "feature")

d145_chi_table <- chi_table
d145_chi_table$comparison <- "d145"
```

combine all chi sq comparisons together
```{r}
all_chi <- rbind(amb_chi_table,d10_chi_table,d135_chi_table,d145_chi_table)

all_chi$feature <- gsub("3prime_UTR", "put.3primeUTR",all_chi$feature)
all_chi$feature <- gsub("put_promoter", "put.promoter",all_chi$feature)
all_chi$feature <- gsub("repeat_region", "repeat.reagion",all_chi$feature)
all_chi$comparison <- gsub("amb", "time", all_chi$comparison)
all_chi$comparison <- gsub("d", "day_", all_chi$comparison)

write.csv(all_chi,"RAnalysis/Output/DMR_genom_feat_chi_table.csv", row.names = FALSE, quote = FALSE)

```


