---
title: "Untitled"
author: "Shelly Trigg"
date: "8/12/2020"
output: html_document
---

load libraries
```{r}
library(data.table)
library(gplots)
library(ggplot2)
library(dplyr)
library(broom)
library(RColorBrewer)
library(egg)
library(purrr)
library(tidyr)
library(ggpubr)
```

# Part 1:  Filter methylated regions for regions that have coverage in 75% of samples per group

read in data
```{r}
DMR <- data.frame(fread("https://gannet.fish.washington.edu/metacarcinus/Salmo_Calig/analyses/20200811/DMR250bp_MCmax25_cov5x_rms_results_collapsed.tsv"))

```



loop through table and keep only lines where up to one sample contains NA for % methylation
```{r}
## For ambient DMR table
df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(DMR))){
  t16_s26 <- DMR[i,7:10] #define columns from 16C_26psu
  t16_s32 <- DMR[i,11:14] #define columns from 16C_32psu
  t8_s26 <- DMR[i,15:18] #define columns from 8C_26psu
  t8_s32 <- DMR[i,19:22] #define columns from 8C_32psu
  if(length(which(is.na(t16_s26))) < 2 & length(which(is.na(t16_s32))) < 2 & length(which(is.na(t8_s26))) < 2 & length(which(is.na(t8_s32))) < 2){
  df <- rbind(df,DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}

```

# Part 2: Run group statistics on regions to find regions that are significantly different among groups

Make a unique ID column 
```{r}
colnames(df)[1] <- "chr"
#for all ambient sample comparison
df$ID <- paste(df$chr,":",df$start,"-",df$end, sep = "")
df$ID <- gsub("__.*__.*:",":",df$ID)

```

reformat data for calculating group effect
```{r}
#reformat all to long format
DMR_STACKED <- tidyr::gather(df[,7:ncol(df)], "Sample.ID", "perc.meth",1:16)

#simplify sample ID column
DMR_STACKED$Sample.ID <- gsub("methylation_level_","", DMR_STACKED$Sample.ID)

#Create treatment column
DMR_STACKED$Treatment <- gsub("psu_.*","psu", DMR_STACKED$Sample.ID)

#create a temperature column
DMR_STACKED$Temp <- gsub("C_.*","",DMR_STACKED$Sample.ID)
DMR_STACKED$Temp <- gsub("CTRL_","", DMR_STACKED$Temp)

#Create a salinity column
DMR_STACKED$Salinity <- gsub(".*C_","", DMR_STACKED$Sample.ID)
DMR_STACKED$Salinity <- gsub("psu_.*", "", DMR_STACKED$Salinity)

#Create an infestation status column
DMR_STACKED$Lice_status <- ifelse(grepl("CTRL",DMR_STACKED$Sample.ID),"no","yes")
```

plot % meth dist.
```{r}
a <- ggplot(DMR_STACKED) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw() + xlab("DMR methylation (# mCpGs/total CpGs )") + theme(text = element_text(size=8),plot.title = element_text(size = 8,face = "bold"),legend.title = element_text(size = 4),legend.text = element_text(size = 4))

```

arc sin sqrt transform the data
```{r}
#arcsin sqrt transformation function
asinTransform <- function(p) { asin(sqrt(p))}

#arcsin transform data 

DMR_STACKED_asin <- DMR_STACKED
DMR_STACKED_asin$perc.meth <- asinTransform(DMR_STACKED_asin$perc.meth)

```

plot distribution of TRANSFORMED % methylation in all DMRs in all samples
```{r}
b <- ggplot(DMR_STACKED_asin) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw() + xlab("transformed DMR methylation (# mCpGs/total CpGs )") + theme(text = element_text(size=8),plot.title = element_text(size = 8,face = "bold"),legend.title = element_text(size = 4),legend.text = element_text(size = 4))

jpeg("DMR_mCpG_histograms.jpg", width = 6, height = 6, units = "in", res = 300)
ggarrange(a,b, nrow = 2)
dev.off()
```

## Run anova on TRANSFORMED data to assess overall group differences
```{r}
# hypothesis:  no effect from treatment
capture.output(summary(aov(perc.meth~Treatment,data = DMR_STACKED_asin)), file = "Results_1way_AOV.txt")

# hypothesis:  no effect from temp, salinity, or their interaction
capture.output(summary(aov(perc.meth~Temp*Salinity,data = DMR_STACKED_asin)), file = "Results_2way_AOV.txt")

```

## Run anova on TRANSFORMED data to assess group differences for each DMR

**Hypotehsis: there is no effect from temp, sal., or their interaction**
```{r}
#
DMR_2way_aov <- DMR_STACKED_asin[which(DMR_STACKED_asin$Lice_status == "yes"),] %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Temp*Salinity, data =  . ))
#summarize ANOVA data
#DMR_2way_aov_modelsumm <- glance(DMR_2way_aov, meth_aov_models)
DMR_2way_aov_modelsumm <- DMR_2way_aov %>% ungroup %>% 
    pull(meth_aov_models) %>% 
    map_dfr(tidy, .id = 'grp')

#spread out pvalues
DMR_2way_aov_modelsumm_wide <- data.frame(tidyr::pivot_wider(DMR_2way_aov_modelsumm, names_from = term, values_from = c("df", "sumsq", "meansq","statistic","p.value"),names_sep = "_" ))




#add group labels back

DMR_2way_aov_modelsumm_wide <- cbind(DMR_2way_aov[,1], DMR_2way_aov_modelsumm_wide[,-c(1)])


#create a new column denoting significant effect
for(i in 1:nrow(DMR_2way_aov_modelsumm_wide)){
  if(DMR_2way_aov_modelsumm_wide$p.value_Temp[i] < 0.05 &  DMR_2way_aov_modelsumm_wide$p.value_Salinity[i] > 0.05 & DMR_2way_aov_modelsumm_wide$p.value_Temp.Salinity[i] > 0.05){
    DMR_2way_aov_modelsumm_wide$sig_effect[i] <- "Temp"
  }
  if(DMR_2way_aov_modelsumm_wide$p.value_Temp[i] > 0.05 &  DMR_2way_aov_modelsumm_wide$p.value_Salinity[i] < 0.05 & DMR_2way_aov_modelsumm_wide$p.value_Temp.Salinity[i] > 0.05){
    DMR_2way_aov_modelsumm_wide$sig_effect[i] <- "Sal"
  }
  if( DMR_2way_aov_modelsumm_wide$p.value_Temp.Salinity[i] < 0.05){
    DMR_2way_aov_modelsumm_wide$sig_effect[i] <- "TempxSal"
  }
  if(DMR_2way_aov_modelsumm_wide$p.value_Temp[i] < 0.05 &  DMR_2way_aov_modelsumm_wide$p.value_Salinity[i] < 0.05 & DMR_2way_aov_modelsumm_wide$p.value_Temp.Salinity[i] > 0.05){
    DMR_2way_aov_modelsumm_wide$sig_effect[i] <- "Temp+Sal"
  }
  if(DMR_2way_aov_modelsumm_wide$p.value_Temp[i] > 0.05 &  DMR_2way_aov_modelsumm_wide$p.value_Salinity[i] > 0.05 & DMR_2way_aov_modelsumm_wide$p.value_Temp.Salinity[i] > 0.05){
    DMR_2way_aov_modelsumm_wide$sig_effect[i] <- "none"
  }
}
 

DMR_2way_aov_modelsumm_wide_0.05 <- DMR_2way_aov_modelsumm_wide[which(DMR_2way_aov_modelsumm_wide$sig_effect!="none"),]


#add other model statistics to the table
DMR_2way_aov_modelsumm_stats <- DMR_2way_aov %>% ungroup %>% 
    pull(meth_aov_models) %>% 
    map_dfr(glance, .id = 'grp')

#add group labels back
DMR_2way_aov_modelsumm_stats <- cbind(DMR_2way_aov[,1], DMR_2way_aov_modelsumm_stats[,-c(1)])


#merge other stats with main stats
DMR_2way_aov_modelsumm_wide_stats <- merge(DMR_2way_aov_modelsumm_stats,DMR_2way_aov_modelsumm_wide, by = "ID")

#add chr, start, and end pos. info to stats table

DMR_2way_aov_modelsumm_wide_stats <- merge(df[,c("chr", "start", "end", "ID")],DMR_2way_aov_modelsumm_wide_stats, by ="ID")

#write out stats table
write.csv(DMR_2way_aov_modelsumm_wide_stats,"DMR_2way_aov_summ_stats.csv", row.names = F, quote = F)


# #run tukey test 
# DMR_2way_aov_tuk <- DMR_STACKED_asin[which(DMR_STACKED_asin$Lice_status == "yes"),] %>% group_by(ID) %>%
# do(meth_tuk_models = TukeyHSD(aov(perc.meth~Temp*Salinity, data =  . )))
# 
# #DMR_2way_aov_tuk_modelsumm <- tidy(DMR_2way_aov_tuk,meth_tuk_models)
# 
# #this works!!!
# #https://stackoverflow.com/questions/62972843/retreiving-tidy-results-from-regression-by-group-with-broom
# DMR_2way_aov_tuk_modelsumm <- DMR_2way_aov_tuk %>% ungroup %>% 
#     pull(meth_tuk_models) %>% 
#     map_dfr(tidy, .id = 'grp')
# 
# DMR_2way_aov_tuk_modelsumm_wide <- data.frame(tidyr::pivot_wider(DMR_2way_aov_tuk_modelsumm, names_from = c("term","comparison"), values_from = c("estimate", "conf.low", "conf.high","adj.p.value"),names_sep = "_" ))
# 
# DMR_2way_aov_tuk_model_termsumm_0.05 <- merge(DMR_2way_aov_model_termsumm_0.05,DMR_2way_aov_tuk_modelsumm_wide, by ="ID")
# 
# #remove tukey data for DMRs that don't have a significant interaction pvalue
# 
# for (i in 1:nrow(DMR_2way_aov_tuk_model_termsumm_0.05)){
#   if(DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp.Salinity[i] > 0.05){
#     DMR_2way_aov_tuk_model_termsumm_0.05[i,33:50] <- "NA"
#   }
# }
# 
# DMR_2way_aov_tuk_model_termsumm_0.05 <- merge(df[,c("ID", "chr","start","end")],DMR_2way_aov_tuk_model_termsumm_0.05,by = "ID" )
# write.csv(DMR_2way_aov_tuk_model_termsumm_0.05,"DMR_2way_aov_tuk_model_summ_0.05.csv", row.names = F, quote = F)
# 
# 
# 
# 
# 


# #filter tuk results for DMRs with overall model pvalue < = 0.05
# DMR_2way_aov0.05_tuk_modelsumm <- DMR_2way_aov_tuk_modelsumm[which(DMR_2way_aov_tuk_modelsumm$ID %in% DMR_2way_aov_modelsumm_0.05$ID),]
# #there are 83 DMRs
# 
# #filter for DMRs with Tuk p <= 0.05
# DMR_2way_aov0.05_tuk_modelsumm_0.05 <- DMR_2way_aov0.05_tuk_modelsumm[which(DMR_2way_aov0.05_tuk_modelsumm$adj.p.value<= 0.05),]
# #there are 81 DMRs; 2 DMRs did not have adj.p.values <= 0.05
# 
# DMR_2way_aov0.05_tuk_modelsumm_0.05$term_comp <- paste(DMR_2way_aov0.05_tuk_modelsumm_0.05$term, DMR_2way_aov0.05_tuk_modelsumm_0.05$comparison, sep = "_")
# 
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread <- data.frame(tidyr::spread(DMR_2way_aov0.05_tuk_modelsumm_0.05[,c(1,7,8)], "term_comp", "adj.p.value"))
# 
# #add column for chr, start and end
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread$chr <- gsub(":.*","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread$start <- gsub(".*:","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread$start <- gsub("-.*","",DMR_2way_aov_tuk_modelsumm_0.05_spread$start)
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread$end <- gsub(".*-","",DMR_2way_aov_tuk_modelsumm_0.05_spread$ID)
# 
# DMR_2way_aov_tuk_modelsumm_0.05_spread_bed <- DMR_2way_aov_tuk_modelsumm_0.05_spread[,c(10:12, 2:9)]
# 
# write.table(DMR_2way_aov_tuk_modelsumm_0.05_spread_bed,"DMR_2way_aov_tuk_modelsumm_0.05.csv", row.names = F, quote = F)
```



## plot heatmap of 2way aov for pH p.val 0.05 sig DMRs
```{r}
#create matrix for heatmap plotting
aov2way_0.05_DMR_m <- as.matrix(df[,7:22])
rownames(aov2way_0.05_DMR_m) <- df$ID

#remove extra text from column names
colnames(aov2way_0.05_DMR_m) <- gsub("methylation_level_","",colnames(aov2way_0.05_DMR_m))

#create matrix for looking at salinity effect
aov2way_0.05_Sal_DMR_m <- aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect == "Sal"),],ID)),]

jpeg("DMR_percMeth_Sal_heatmap.jpg", width = 8, height = 10, units = "in", res = 300)
heatmap.2(aov2way_0.05_Sal_DMR_m,margins = c(10,20), cexRow = 0.75,cexCol = 1.25,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_Sal_DMR_m),rowsep=1:nrow(aov2way_0.05_Sal_DMR_m),lhei = c(1,7), ColSideColors = c(rep("magenta4",4), rep("magenta2",4), rep("cyan4",4), rep("cyan2",4)))
dev.off()

#create matrix for looking at temp effect
aov2way_0.05_temp_DMR_m <-  aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect == "Temp"),],ID)),]

jpeg("DMR_percMeth_Temp_heatmap.jpg", width = 8, height = 10, units = "in", res = 300)
heatmap.2(aov2way_0.05_temp_DMR_m, margins = c(15,20),cexRow = 1,cexCol = 1.25,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_temp_DMR_m),rowsep=1:nrow(aov2way_0.05_temp_DMR_m),lhei = c(1,7),ColSideColors = c(rep("magenta4",4), rep("magenta2",4), rep("cyan4",4), rep("cyan2",4)))
dev.off()

#create matrix for looking at interaction effect
aov2way_0.05_tempxsal_DMR_m <-  aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect== "TempxSal"),],ID)),]

jpeg("DMR_percMethTempXSal_heatmap.jpg", width = 8, height = 10, units = "in", res = 300)
heatmap.2(aov2way_0.05_tempxsal_DMR_m,margins = c(15,20), cexRow = 1,cexCol = 1.25,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_tempxsal_DMR_m),rowsep=1:nrow(aov2way_0.05_tempxsal_DMR_m),lhei = c(1,7),ColSideColors = c(rep("magenta4",4), rep("magenta2",4), rep("cyan4",4), rep("cyan2",4)))
dev.off()


#plot DMRs with additive effect

#create matrix for looking at interaction effect
aov2way_0.05_tempANDsal_DMR_m <-  aov2way_0.05_DMR_m[which(rownames(aov2way_0.05_DMR_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect== "Temp+Sal"),],ID)),]

jpeg("DMR_percMethTempANDSal_heatmap.jpg", width = 12, height = 10, units = "in", res = 300)
heatmap.2(aov2way_0.05_tempANDsal_DMR_m, margins = c(15,20),cexRow = 1.25, cexCol = 1.5,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_tempANDsal_DMR_m),rowsep=1:nrow(aov2way_0.05_tempANDsal_DMR_m),lhei = c(1,7),ColSideColors = c(rep("magenta4",4), rep("magenta2",4), rep("cyan4",4), rep("cyan2",4)))
dev.off()

# #create multipanel plot of heatmaps
# 
# library(gridGraphics)
# library(grid)
# 
# grab_grob <- function(){
#   grid.echo()
#   grid.grab()
# }
# 
# arr <- list(aov2way_0.05_temp_DMR_m,aov2way_0.05_Sal_DMR_m, aov2way_0.05_tempxsal_DMR_m, aov2way_0.05_tempANDsal_DMR_m)
# 
# gl <- lapply(1:4, function(i){
#   heatmap.2(arr[[i]], margins = c(10,20), cexRow = 0.2,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", dendrogram = "row", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(arr[[i]]),rowsep=1:nrow(arr[[i]])          )
#   grab_grob()
# })
# 
# grid.newpage()
# library(gridExtra)
# jpeg("DMR_sig0.05_heatmaps.jpg", width = 12, height = 6, units = "in", res = 300)
# grid.arrange(grobs=gl, ncol=4, clip=TRUE)
# dev.off()

```

###Visualize group means
```{r}
#calculate group means
MeanT16S26 <- rowMeans(aov2way_0.05_DMR_m[,1:4], na.rm = TRUE)
MeanT16S32 <- rowMeans(aov2way_0.05_DMR_m[,5:8], na.rm = TRUE)
MeanT8S26 <- rowMeans(aov2way_0.05_DMR_m[,9:12], na.rm = TRUE)
MeanT8S32 <- rowMeans(aov2way_0.05_DMR_m[,13:16], na.rm = TRUE)


#bind all group means together
aov2way_0.05_DMR_mean_m <- as.matrix(data.frame(cbind(MeanT16S26, MeanT16S32, MeanT8S26, MeanT8S32)))

#plot for day10 group mean comparison
colnames(aov2way_0.05_DMR_mean_m) <- c("16C_26psu", "16C_32psu", "8C_26psu", "8C_32psu")


#plot DMRs that show temerature effect @ p < 0.05
aov2way_0.05_temp_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect=="Temp"),],ID)),]

jpeg("DMR_meanMeth_Temp_heatmap.jpg", width = 8, height = 10,units = "in", res = 300 )
heatmap.2(aov2way_0.05_temp_DMR_mean_m,margins = c(7,20), cexCol = 1,cexRow = 0.8,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_temp_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_temp_DMR_mean_m),lhei = c(1,7))
dev.off()

#plot DMR means that show salinity effect @ p < 0.05
aov2way_0.05_sal_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect =="Sal"),],ID)),]

jpeg("DMR_meanMeth_Sal_heatmap.jpg", width = 8, height = 10,units = "in", res = 300 )
heatmap.2(aov2way_0.05_sal_DMR_mean_m,margins = c(7,20), cexCol = 1, cexRow = 0.75, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_sal_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_sal_DMR_mean_m),lhei = c(1,7))
dev.off()

#plot DMRs that show interaction effect @ p < 0.05
aov2way_0.05_salxtemp_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect=="TempxSal"),],ID)),]

jpeg("DMR_meanMeth_TempxSal_heatmap.jpg", width = 8, height = 10,units = "in", res = 300)
heatmap.2(aov2way_0.05_salxtemp_DMR_mean_m,margins = c(7,20), cexCol = 1,cexRow = 1, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_salxtemp_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_salxtemp_DMR_mean_m),lhei = c(1,7))
dev.off()

#plot DMRs that show additive effect @ p < 0.05
aov2way_0.05_salANDtemp_DMR_mean_m <- aov2way_0.05_DMR_mean_m[which(rownames(aov2way_0.05_DMR_mean_m) %in% pull(DMR_2way_aov_modelsumm_wide_0.05[which(DMR_2way_aov_modelsumm_wide_0.05$sig_effect=="Temp+Sal"),],ID)),]

jpeg("DMR_meanMeth_TempANDSal_heatmap.jpg", width = 8, height = 10,units = "in", res = 300)
heatmap.2(aov2way_0.05_salANDtemp_DMR_mean_m,margins = c(7,15), cexCol = 1, cexRow = 1,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),dendrogram = "row",Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov2way_0.05_salANDtemp_DMR_mean_m),rowsep=1:nrow(aov2way_0.05_salANDtemp_DMR_mean_m),lhei = c(1,7))
dev.off()

```



#Create dfs for exclusively salinity effect, temp effect, and interaction effect
```{r}
#based on Tukey's HSD p. values
DMR_2way_aov_0.05_temp <- DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp_8.16 < 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Salinity_32.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.26.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_16.32.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_16.32.8.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.8.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.16.32 > 0.05),]
# there is only one DMR with a significant temperature term and no significant interaction term

DMR_2way_aov_0.05_sal <- DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Salinity_32.26 < 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp_8.16 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.26.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_16.32.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.16.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_16.32.8.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.8.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp.Salinity_8.32.16.32 > 0.05),]
# there are only 2 DMR with a significant salinity term and no significant interaction term
DMR_2way_aov_0.05_salxtemp <-  DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Salinity_32.26 > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$adj.p.value_Temp_8.16 > 0.05),]  
#there are 16 DMRs that have significant interaction terms and not significant salinity or temperature terms

# based on ANOVA effect p.values

DMR_2way_aov_0.05_temp <- DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp < 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Salinity > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp.Salinity > 0.05),]
# there are 11 DMRs with a significant temperature term and no significant interaction term

DMR_2way_aov_0.05_sal <- DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Salinity < 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp.Salinity > 0.05),]
# there are 33 DMRs with a significant salinity term and no significant interaction term

DMR_2way_aov_0.05_salxtemp <- DMR_2way_aov_tuk_model_termsumm_0.05[which(DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Salinity > 0.05 & DMR_2way_aov_tuk_model_termsumm_0.05$p.value_Temp.Salinity < 0.05),]
# there are 15 DMRs with a significant interaction term and no significant terms

```
